---
title: "Take Home Ex 2"
date: "20 Feburary 2023"
date-modified: "`r Sys.Date()`"
format:
 html:
  toc: true
  toc-location: right
  number-depth: 3
execute: 
  message: false
  warning: false
editor: visual
---

```{r}
pacman::p_load(dplyr, sf, spdep, tmap, tidyverse, readxl, ggplot2,moments,plotly, sfdep,Kendall, spacetime, lubridate,anytime)
```

```{r}
jul<- read_excel("data/aspatial/Juli 2021.xlsx")
aug21<- read_excel("data/aspatial/Agustus 2021.xlsx")
sep21<- read_excel("data/aspatial/September 2021.xlsx")
oct21<- read_excel("data/aspatial/Oktober 2021.xlsx")
nov21<- read_excel("data/aspatial/November 2021.xlsx")
dec21<- read_excel("data/aspatial/Desember 2021.xlsx")
jan22<- read_excel("data/aspatial/Januari 2022.xlsx")
feb22<- read_excel("data/aspatial/Februari 2022.xlsx")
mar22<- read_excel("data/aspatial/Maret 2022.xlsx")
apr22<- read_excel("data/aspatial/April 2022.xlsx")
may22<- read_excel("data/aspatial/Mei 2022.xlsx")
jun22<- read_excel("data/aspatial/Juni 2022.xlsx")
```

```{r}
colnames(jul)[6] ="NotYetV_jul21"
colnames(jul)[7] ="D1_jul21"
colnames(jul)[8] ="D2_jul21"
colnames(jul)[9] ="TotalV_jul21"
jul21<-jul[-c(1:3,5,10:27)]


colnames(aug21)[6] ="NotYetV_aug21"
colnames(aug21)[7] ="D1_aug21"
colnames(aug21)[8] ="D2_aug21"
colnames(aug21)[9] ="TotalV_aug21"
aug21<-aug21[-c(1:3,5,10:27)]

colnames(sep21)[6] ="NotYetV_sep21"
colnames(sep21)[7] ="D1_sep21"
colnames(sep21)[8] ="D2_sep21"
colnames(sep21)[9] ="TotalV_sep21"
sep21<- sep21[-c(1:3,5,10:27)]

colnames(oct21)[6] ="NotYetV_oct21"
colnames(oct21)[7] ="D1_oct21"
colnames(oct21)[8] ="D2_oct21"
colnames(oct21)[9] ="TotalV_oct21"
oct21<-oct21[-c(1:3,5,10:27)]

colnames(nov21)[6] ="NotYetV_nov21"
colnames(nov21)[7] ="D1_nov21"
colnames(nov21)[8] ="D2_nov21"
colnames(nov21)[9] ="TotalV_nov21"
nov21<-nov21[-c(1:3,5,10:27)]

colnames(dec21)[6] ="NotYetV_dec21"
colnames(dec21)[7] ="D1_dec21"
colnames(dec21)[8] ="D2_dec21"
colnames(dec21)[9] ="TotalV_dec21"
dec21<-dec21[-c(1:3,5,10:27)]

colnames(jan22)[6] ="NotYetV_jan21"
colnames(jan22)[7] ="D1_jan22"
colnames(jan22)[8] ="D2_jan22"
colnames(jan22)[9] ="TotalV_jan22"
jan22<-jan22[-c(1:3,5,10:27)]

colnames(feb22)[6] ="NotYetV_jan21"
colnames(feb22)[7] ="D1_feb22"
colnames(feb22)[8] ="D2_feb22"
colnames(feb22)[9] ="TotalV_feb22"
feb22<-feb22[-c(1:3,5,10:27)]

colnames(mar22)[6] ="NotYetV_mar21"
colnames(mar22)[7] ="D1_mar22"
colnames(mar22)[8] ="D2_mar22"
colnames(mar22)[9] ="D3_mar22"
colnames(mar22)[10] ="TotalV_mar22"
mar22<-mar22[-c(1:3,5,11:34)]

colnames(apr22)[6] ="NotYetV_apr21"
colnames(apr22)[7] ="D1_apr22"
colnames(apr22)[8] ="D2_apr22"
colnames(apr22)[9] ="D3_apr22"
colnames(apr22)[10] ="TotalV_apr22"
apr22<-apr22[-c(1:3,5,11:34)]

colnames(may22)[6] ="NotYetV_may21"
colnames(may22)[7] ="D1_may22"
colnames(may22)[8] ="D2_may22"
colnames(may22)[9] ="D3_may22"
colnames(may22)[10] ="TotalV_may22"
may22<-may22[-c(1:3,5,11:34)]

colnames(jun22)[6] ="NotYetV_jun21"
colnames(jun22)[7] ="D1_jun22"
colnames(jun22)[8] ="D2_jun22"
colnames(jun22)[9] ="D3_jun22"
colnames(jun22)[10] ="TotalV_jun22"
jun22<-jun22[-c(1:3,5,11:34)]
```

```{r}
FY21_22Vaccine <-jul[c(1:9)]
FY21_22Vaccine<- left_join(FY21_22Vaccine, aug21, by="KELURAHAN")
FY21_22Vaccine<- left_join(FY21_22Vaccine, sep21, by="KELURAHAN")
FY21_22Vaccine<- left_join(FY21_22Vaccine, oct21, by="KELURAHAN")
FY21_22Vaccine<- left_join(FY21_22Vaccine, nov21, by="KELURAHAN")
FY21_22Vaccine<- left_join(FY21_22Vaccine, dec21, by="KELURAHAN")
FY21_22Vaccine<- left_join(FY21_22Vaccine, jan22, by="KELURAHAN")
FY21_22Vaccine<- left_join(FY21_22Vaccine, feb22, by="KELURAHAN")
FY21_22Vaccine<- left_join(FY21_22Vaccine, mar22, by="KELURAHAN")
FY21_22Vaccine<- left_join(FY21_22Vaccine, apr22, by="KELURAHAN")
FY21_22Vaccine<- left_join(FY21_22Vaccine, may22, by="KELURAHAN")
FY21_22Vaccine<- left_join(FY21_22Vaccine, jun22, by="KELURAHAN")

```

```{r}
jakarta <- st_read(dsn = "data/geospatial", 
                layer = "BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA") %>% st_transform(crs=23845)

jakarta <- jakarta[c(1:9)]
```

```{r}
st_crs(jakarta)
```

```{r}
qtm(jakarta)
```

```{r}
jakarta<- filter(jakarta, KAB_KOTA != "KEPULAUAN SERIBU")
```

```{r}
qtm(jakarta)
```

```{r}
jakarta_vaccine <- left_join(jakarta, FY21_22Vaccine,
                          by = c("KODE_DESA" = "KODE KELURAHAN"))
```

```{r}
jakarta_vaccine <- st_as_sf(jakarta_vaccine)
```

```{r}
tmap_mode("plot")
qtm(jakarta_vaccine, 
    fill = "KAB_KOTA"
    )
```

```{r}
tmap_mode("view")

qtm(jakarta_vaccine, 
    fill = "JUMLAH_PEN",
        text= "KELURAHAN",
    text.size = 0.5)

tmap_mode("plot")
```

```{r}
jul21total = data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month =  ymd("2021-07-31"),
                           TotalDose = jakarta_vaccine$TotalV_jul21,
                          population = jakarta_vaccine$JUMLAH_PEN,
                           geometry = jakarta_vaccine$geometry
                           )

aug21total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2021-08-31"),
                           TotalDose = jakarta_vaccine$TotalV_aug21,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)
sep21total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2021-09-30"),
                           TotalDose = jakarta_vaccine$TotalV_sep21,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)
oct21total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2021-10-31"),
                           TotalDose = jakarta_vaccine$TotalV_oct21,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)
nov21total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2021-11-30"),
                           TotalDose = jakarta_vaccine$TotalV_nov21,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)
dec21total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2021-12-31"),
                           TotalDose = jakarta_vaccine$TotalV_dec21,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)

jan22total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2022-01-31"),
                           TotalDose = jakarta_vaccine$TotalV_jan22,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)
feb22total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2022-02-28"),
                           TotalDose = jakarta_vaccine$TotalV_feb22,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)
mar22total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2022-03-31"),
                           TotalDose = jakarta_vaccine$TotalV_mar22,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)
apr22total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2022-04-30"),
                           TotalDose = jakarta_vaccine$TotalV_apr22,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)
may22total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2022-05-31"),
                           TotalDose = jakarta_vaccine$TotalV_may22,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)
jun22total <- data.frame(SubDistrict = jakarta_vaccine$KELURAHAN,
                            Month = ymd("2022-06-30"),
                           TotalDose = jakarta_vaccine$TotalV_jun22,
                         population = jakarta_vaccine$JUMLAH_PEN,
                         geometry = jakarta_vaccine$geometry)
```

```{r}
total_vaccine <- rbind(jul21total, aug21total,sep21total,oct21total,nov21total,dec21total,jan22total,feb22total,mar22total,apr22total,may22total,jun22total)
```

```{r}
jenks_plot <- function(df, varname) {
  tm_shape(jakarta_vaccine) +
    tm_polygons() +
  tm_shape(df) +
    tm_fill(varname, 
          n= 6,
          style = "jenks", 
          title = "No of people vaccinated") +
    tm_layout(main.title = varname,
          main.title.position = "center",
          main.title.size = 1.2,
          legend.height = 0.45, 
          legend.width = 0.35,
          frame = TRUE) +
    tm_borders(alpha = 0.2)
}
```

```{r}
tmap_mode("plot")
tmap_arrange(jenks_plot(jakarta_vaccine, "D1_jul21"),
             jenks_plot(jakarta_vaccine, "D2_jul21"),
             jenks_plot(jakarta_vaccine, "D1_aug21"),
             jenks_plot(jakarta_vaccine, "D2_aug21"))
```

```{r}
tmap_arrange(jenks_plot(jakarta_vaccine, "D1_sep21"),
             jenks_plot(jakarta_vaccine, "D2_sep21"),
             jenks_plot(jakarta_vaccine, "D1_oct21"),
             jenks_plot(jakarta_vaccine, "D2_oct21"))
```

```{r}
tmap_arrange(jenks_plot(jakarta_vaccine, "D1_nov21"),
             jenks_plot(jakarta_vaccine, "D2_nov21"),
             jenks_plot(jakarta_vaccine, "D1_dec21"),
             jenks_plot(jakarta_vaccine, "D2_dec21"))
```

```{r}
tmap_arrange(jenks_plot(jakarta_vaccine, "D1_jan22"),
             jenks_plot(jakarta_vaccine, "D2_jan22"),
             jenks_plot(jakarta_vaccine, "D1_feb22"),
             jenks_plot(jakarta_vaccine, "D2_feb22"))
```

```{r}
tmap_arrange(jenks_plot(jakarta_vaccine, "D1_mar22"),
             jenks_plot(jakarta_vaccine, "D2_mar22"),
             jenks_plot(jakarta_vaccine, "D1_apr22"),
             jenks_plot(jakarta_vaccine, "D2_apr22"))
```

```{r}
tmap_arrange(jenks_plot(jakarta_vaccine, "D1_may22"),
             jenks_plot(jakarta_vaccine, "D2_may22"),
             jenks_plot(jakarta_vaccine, "D1_jun22"),
             jenks_plot(jakarta_vaccine, "D2_jun22"))
```

```{r}
tmap_mode("plot")
tmap_arrange(jenks_plot(jakarta_vaccine, "TotalV_jul21"),
             jenks_plot(jakarta_vaccine, "TotalV_aug21"),
             jenks_plot(jakarta_vaccine, "TotalV_sep21"),
             jenks_plot(jakarta_vaccine, "TotalV_oct21"))
```

```{r}
tmap_arrange(jenks_plot(jakarta_vaccine, "TotalV_nov21"),
             jenks_plot(jakarta_vaccine, "TotalV_dec21"),
             jenks_plot(jakarta_vaccine, "TotalV_jan22"),
             jenks_plot(jakarta_vaccine, "TotalV_feb22"))
```

```{r}
tmap_arrange(jenks_plot(jakarta_vaccine, "TotalV_mar22"),
             jenks_plot(jakarta_vaccine, "TotalV_apr22"),
             jenks_plot(jakarta_vaccine, "TotalV_may22"),
             jenks_plot(jakarta_vaccine, "TotalV_jun22"))
```

## Computing Local Gi statistics

Unlike global measures that summarize the overall spatial autocorrelation of the study area in one single value, local measures of spatial association identify local clusters (observations nearby have similar attribute values) or spatial outliers (observations nearby have different attribute values).

In the below section we will then check if we can find the sub district in which the \*\*relatively higher number of vaccination\*\* rate is significantly higher.

we will be using local_gstar_perm() functions to compute the Local Gi statistic.

Local Gi stats allows the detection of a local concentration of high and low values in neighboring objects and studies the statistical significance of that dependence.

```{r}
longitude <- map_dbl(jakarta$geometry, ~st_centroid(.x)[[1]])
latitude <- map_dbl(jakarta$geometry, ~st_centroid(.x)[[2]])
coords <- cbind(longitude, latitude)
```

```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
all.linked <- max(unlist(nbdists(k1, coords, longlat = TRUE)))
summary(k1dists)
```

```{r}
fix_d <- dnearneigh(coords, 0, all.linked, longlat = TRUE)
fix_d
```

```{r}
fix_lw <- nb2listw(fix_d, style = 'W')
summary(fix_lw)
```

```{r}
set.seed(999)
localgi_jul21 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_aug21 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_sep21 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_oct21 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_nov21 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_dec21 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_jan22 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_feb22 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_mar22 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_apr22 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_may22 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)
localgi_jun22 <- local_gstar_perm(jakarta_vaccine$TotalV_jul21, fix_d, fix_lw, nsim=39)

```

```{r}
jv.localGi_jul21 <- cbind(jakarta_vaccine,localgi_jul21)
jv.localGi_aug21 <- cbind(jakarta_vaccine,localgi_aug21)
jv.localGi_sep21 <- cbind(jakarta_vaccine,localgi_sep21)
jv.localGi_oct21 <- cbind(jakarta_vaccine,localgi_oct21)
jv.localGi_nov21 <- cbind(jakarta_vaccine,localgi_nov21)
jv.localGi_dec21 <- cbind(jakarta_vaccine,localgi_dec21)
jv.localGi_jan22 <- cbind(jakarta_vaccine,localgi_jan22)
jv.localGi_feb22 <- cbind(jakarta_vaccine,localgi_feb22)
jv.localGi_mar22 <- cbind(jakarta_vaccine,localgi_mar22)
jv.localGi_apr22 <- cbind(jakarta_vaccine,localgi_apr22)
jv.localGi_may22 <- cbind(jakarta_vaccine,localgi_may22)
jv.localGi_jun22 <- cbind(jakarta_vaccine,localgi_jun22)
```

```{r}
tmap_mode("plot")
tm_shape(jv.localGi_jun22) +
    tm_polygons() +
    tm_shape(jv.localGi_jun22 %>% filter(p_value < 0.05)) +
    tm_fill(col= "p_sim",
             palette = "Paired") +
    tm_borders(alpha = 0.4) +
    tm_layout(main.title = paste("local Gi p-value <0.05 JUN22"),
              main.title.size = 1)
```

```{r}
localgi_HCplot <- function(df, varname) {
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df %>% filter(p_value < 0.05)) +
    tm_fill(varname,
            palette = "-RdBu") +
    tm_layout(
          legend.height = 0.45, 
          legend.width = 0.35,
          frame = TRUE) +
    tm_borders(alpha = 0.2)
}
```

```{r}
tmap_arrange(localgi_HCplot(jv.localGi_jul21, "gi_star"),
             localgi_HCplot(jv.localGi_aug21, "gi_star"),
             localgi_HCplot(jv.localGi_sep21, "gi_star"),
             localgi_HCplot(jv.localGi_oct21, "gi_star")
             )
```

```{r}
jun2022testing <- jv.localGi_jun22 %>% filter(p_value < 0.05)
```
