---
title: "In-Class Ex04"
date: "30 January 2023"
date-modified: "`r Sys.Date()`"
format: html
execute: 
  echo: true
  eval: true
  message: false
  warning: false
editor: visual
---

## Installing and Loading R package

-   spatstat is mainly for spatial point analytics

```{r}
pacman::p_load(maptools, sf, raster, spatstat, tmap)
```

### Importing the spatial data

-   data are in meters and are in ***object format***

Childcare is a single file

```{r}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```

Read in list of data files in the folder

```{r}
sg_sf <- st_read(dsn = "data", layer="CostalOutline")
```

```{r}
mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots(alph=0.5,
          size= 0.01) + 
  tm_view(set.zoom.limits = c(11,14))
```

-   tmap_mode('view') will always connect to the server. it is best not to include more than 5.

-   tm_bubbles create proportional symbol maps.

-   tm_view(set.zoom.limits = c(11,14)):

    -   (a,b) a is the zoom out value, b is the zoom in

# 3 steps to take note

1.  The data table is in object format.

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

2.  It drops the object and display it as a polygon format.

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

3.  this will show a list of the data points.

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```
